Untitled

### 2022-05-31 Manuscript/Lanuza_et_al_2021.Rmd
edits

Trait missing values were imputed with the function *missForest* [@stekhoven2012] which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of *missForest* [@penone2014]. To extract the eigenvectors, we used the function *PVRdecomp* from the package *PVR* [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species, 5.79% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01% of missing values).

@@

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, *P* \< 0.01 and Pearson's correlation = 0.36, *P* \< 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers [@serneels2008; @legendre2012], we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range as indicated for principal component analysis [@legendre2012]. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation with the help of an Anova and a Tukey test. We performed the pPCA using the function *phyl.pca* from the package *phytools* [@revell2012] with the method lambda ($\lambda$) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation [@abdi2010]. Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S3). We found little difference between the explained variance with the imputed dataset (51.08%) and the dataset without missing values (52.87%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

>>>>

To avoid the loss of relevant ecological information that can lead to invalid inferences [@kleinke2020], we imputed trait missing values with the help of the function *missForest* [@stekhoven2012] which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of *missForest* [@penone2014]. To extract the eigenvectors, we used the function *PVRdecomp* from the package *PVR* [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species, 5.79% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without missing values and we found consistent similar results on both analyses (Fig S3 and Fig S4).

[Esta última parte intenta justificar la validación de nuestros resultados, le vendría bien ojo crítico, antes estaba en el siguiente párrafo pero la maldad mental para criticarnos se genera en este y pienso que ya no llega al siguiente con buenos ojos](Jose)

@@

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, *P* \< 0.01 and Pearson's correlation = 0.36, *P* \< 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers [@serneels2008; @legendre2012], we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range as indicated for principal component analysis [@legendre2012]. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation with the help of an Anova and a Tukey test. We performed the pPCA using the function *phyl.pca* from the package *phytools* [@revell2012] with the method lambda ($\lambda$) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation [@abdi2010]. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

### 2022-05-31 Manuscript/Lanuza_et_al_2021.tex
edits

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation. For this, we first implemented a linear model (one for each principal component) where the principal components were the response variable and the different qualitative variables the predictors. Finally, we explore the statistical association with an Anova and the differences between categories for each categorical variable with a Tukey test. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre, 2012}; \protect\hyperlink{ref-serneels2008}{Serneels \& Verdonck, 2008}), we excluded values within the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell, 2012}) with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation (\protect\hyperlink{ref-abdi2010}{Abdi \& Williams, 2010}). Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S3). We found little difference between the explained variance with the imputed dataset (51.08\%) and the dataset without missing values (52.87\%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

>>>>

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre, 2012}; \protect\hyperlink{ref-serneels2008}{Serneels \& Verdonck, 2008}), we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range as indicated for principal component analysis (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre, 2012}). Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation with the help of an Anova and a Tukey test. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell, 2012}) with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation (\protect\hyperlink{ref-abdi2010}{Abdi \& Williams, 2010}). Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S3). We found little difference between the explained variance with the imputed dataset (51.08\%) and the dataset without missing values (52.87\%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

### 2022-05-31 Supplementary_Material/trial.Rmd
edits

---
title: "Untitled"
output: html_document
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10, fig.cap="\\textbf{Fig. S6.} Phylogenetic informed principal component analysis for the subset of species with floral rewards."}
setwd("~/R_Projects/Reproductive Traits")

@@

#Load data
phyl_pca_forest_nectar_pollen <- readRDS("Data/RData/phyl_pca_forest_nectar_pollen.rds")
dat_cleaning_7 <- readRDS("Data/RData/data_all_species_PPCA_nectar_pollen.rds")

@@

library(magick)

#Images were made by Paula Martin
PC$S[,1] <- -PC$S[,1]

@@

  plot <- plot + geom_point(data=dat, aes(x, y),size=1.3, color="black")

  ########
  #ADD ARROWS 
  ########
  datapc <- data.frame(PC$L) #CREATE DATAFRAME WITH LOADINGS
  mult <- min(
    (max(data[,y]) - min(data[,y])/(max(datapc[,y])-min(datapc[,y]))),
    (max(data[,x]) - min(data[,x])/(max(datapc[,x])-min(datapc[,x])))
  )
  datapc <- transform(datapc,
                      v1 = .7 * mult * (get(x)),
                      v2 = .7 * mult * (get(y))
  )
  # add arrows
  plot <- plot + geom_segment(data=datapc,linejoin="round", lineend="round",aes(x=0, y=0, xend=-v1, yend=-v2),size=1.8, arrow=arrow(length=unit(0.5,"cm")), alpha=0.8, colour=c("black"))

  #ADD THE OTHER DIRECTION OF THE SEGMENT BECAUSE LOOKS COOL
  plot <- plot + geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),size=1.6, arrow=arrow(length=unit(0,"cm")),linetype=2, alpha=0.5, colour=c("black"))

@@

  PCAloadings <- data.frame(Variables = rownames(PC$L), PC$L)
  plot <- plot + annotate("text", x = -(PCAloadings$PC1*c(4.85,4.8,4.8,5.5,5,5.5,5.8,6.5)), y = -(PCAloadings$PC2*c(4.8,4.4,4.15,5.35,5,4.7,5.75,6)+c(0,0,0,0,0,0,0,0)),
                          label = PCAloadings$Variables, color="black",fontface =2,size=4.5)

  plot <- plot + theme_ms() +ylim(-4,4) + xlim(-4,4) +  theme(legend.position = c(0.095, 0.11)) +ggtitle("")

>>>>

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,  fig.cap="\\textbf{Fig. S3.}  Phylogenetic informed principal component analysis for the species that did not have missing values."}
setwd("~/R_Projects/Reproductive Traits")

@@

#Load data
phyl_pca_forest_nectar_pollen <- readRDS("Data/RData/phyl_pca_forest_nectar_pollen_complete_cases.rds")
dat_cleaning_5 <- readRDS("Data/RData/data_all_species_PPCA_nectar_pollen_complete_cases.rds")

@@

#############################################################################################################################################################
#COMPATIBILITY
#############################################################################################################################################################
#Load plots

@@

  plot <- plot + geom_point(data=dat, aes(x, y),size=0.95, color="black")

  ########
  #ADD ARROWS 
  ########
  datapc <- data.frame(PC$L) #CREATE DATAFRAME WITH LOADINGS
  mult <- min(
    (max(data[,y]) - min(data[,y])/(max(datapc[,y])-min(datapc[,y]))),
    (max(data[,x]) - min(data[,x])/(max(datapc[,x])-min(datapc[,x])))
  )
  datapc <- transform(datapc,
                      v1 = .7 * mult * (get(x)),
                      v2 = .7 * mult * (get(y))
  )
  # add arrows
  plot <- plot + geom_segment(data=datapc,linejoin="round", lineend="round",aes(x=0, y=0, xend=v1, yend=-v2),size=1.8, arrow=arrow(length=unit(0.5,"cm")), alpha=0.8, colour=c("black"))

  #ADD THE OTHER DIRECTION OF THE SEGMENT BECAUSE LOOKS COOL
  plot <- plot + geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=-v2),size=1.6, arrow=arrow(length=unit(0,"cm")),linetype=2, alpha=0.5, colour=c("black"))

@@

  PCAloadings <- data.frame(Variables = rownames(PC$L), PC$L)
  plot <- plot + annotate("text", x = (PCAloadings$PC1*c(4.6,4.95,5.5,6.25,6.3,4.2,6,6)), y = -(PCAloadings$PC2*c(4.693,3.5,3,6.2,3.5,4.15,6,7.5)+c(0,0,0,0,2,0,0,0)),
                          label = PCAloadings$Variables, color="black",fontface =2,size=4)

  plot <- plot + theme_bw() +ylim(-4,4) + xlim(-4,4) +  theme(legend.position = c(0.095, 0.11)) +ggtitle("")
